Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-27 cvx-flats-neurips21/cvx-flats.tex
Fixing Figure 3
Commenting out G.1

        \begin{scope}
          \clip (h) -- (m) -- (l) -- (h);
          \foreach \c in {0,0.1,...,1}
          {
            \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
            \pgfmathsetmacro\bb{0.5*\c}
            \coordinate (A) at ($(\aa,1-\aa,0)$);
            \coordinate (B) at ($(1-\aa,\aa,0)$);
            \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
            \draw[myblue,thin] (C) parabola (A);
            \draw[myblue,thin] (C) parabola (B);
          }
          \def\c{0.6}
          \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
          \pgfmathsetmacro\bb{0.5*\c}
          \coordinate (A) at ($(\aa,1-\aa,0)$);
          \coordinate (B) at ($(1-\aa,\aa,0)$);
          \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
          \coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
          \coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
          \coordinate (centerdot) at (0.21,0.445, 0.345);
          \draw[myblue, very thick] (C) parabola (A);
          \draw[myblue, very thick] (C) parabola (B);
          \draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
          \node at (centerdot) {$\bullet$};
        \end{scope}
        \def\c{0.6}
        \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
        \pgfmathsetmacro\bb{0.5*\c}
        \node[inner sep=2pt] (levset1) at (0.45,0.55, -0.3) {\vphantom{$\Gamma_{0.1}$}};
        \node (levset) at (0.5,0.5,-.15) {$\Gamma_{0.6}$};
        \draw[-latex,thin] (levset) edge[in=330,out=180] ($(\aa,1-\aa,0)$);
        \draw[-latex,thin] (levset) edge[in=210,out=0] ($(1-\aa,\aa,0)$);
        \placefiglabel{(Q4)}
      \end{tikzpicture}
      \hspace*{-10pt}
      % \\
      % \hline
  \end{tabular}
\caption{Example properties for each quadrant. 
Throughout, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
(Q1,Q2) We cannot fit a $1$-flat (line) through $p$ without leaving the level sets $\gamma_\bot$ and $\gamma_{\text{median}}$, respectively;
Theorem~\ref{thm:convex-flats-inf-dim} implies that there is no $1$-dimensional consistent convex surrogate for either problem.
(Q3) Squared error is a $1$-dimensional convex loss, and indeed it elicits the mean of $Y$, whose level sets are all $1$-flats.
(Q4) The level sets of the variance are curved and cannot fit a $1$-flat; from Theorem~\ref{thm:convex-flats-inf-dim} there is no $1$-dimensional convex surrogate consistent for the variance.}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

Using our $d$-flats framework, we resolve both open questions with new lower bounds in both settings.
% The central observation is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are
Our framework clarifies the relationship between $d$-identifiable properties and properties elicited by $d$-dimensional convex losses: the level sets of the former are $d$-flats by definition, while the level sets of the latter are \emph{unions} of $d$-flats by Theorem~\ref{thm:convex-flats-inf-dim}.
A careful examination of the arguments of~\citet{frongillo2020elicitation} and~\citet{dearborn2020indirect} reveals that they largely rely on the containment of $d$-flats in level sets, rather than the full structure of identifiable properties.
As such, although quite subtle in the case of risk measures, the general structure of these previous proofs go through for convex-elicitable properties: since no $d$-flat could be contained in a particular level set, no union of $d$-flats could be either.
Our lower bounds therefore match both of these papers, though we conjecture that the convex consistency dimension is strictly higher in some cases.

@@

The problem of estimating a risk or uncertainty measure of $Y$ is of central importance in financial regulation~\cite{acerbi2014back-testing,fissler2016expected,campbell2005review} and robust engineering design\raft{add cites from AFOSR or EnvAI}.
% , and algorithmic fairness~\citep{williamson2019fairness}
Risk measures include the upper confidence bound $\E[Y] + \lambda \sqrt{\Var[Y]}$, or the Conditional Value at Risk (CVaR) defined below, in either conditional or unconditional contexts.
Uncertainty measures include the variance, entropy, or norm of the distribution of $Y$.
Risk and uncertainty measures are typically not elicitable, so this problem falls under Quadrant 4.
\citet{frongillo2020elicitation} give prediction dimension lower bounds for a broad class of risk and uncertainty measures, namely Bayes risks.
As stated above, these bounds are with respect to identifiable properties, and bounds for convex surrogates are left as a major open question.

@@

\section{Omitted Examples}\label{app:omitted-examples}
\subsection{Risk-averse classification (Q2)}

Recall the risk-averse classification from Figure~\ref{fig:quadrants-simplices} (Q2).
We can bound Convex Consistency Dimension of this classification task by Corollary~\ref{cor:Pcodim-flat-single-val-prop} on the distribution $p = (0.3,0.3,0.4)$, which we can apply since $\gamma(p) = \{$snowy$\}$, suggest that the convex elicitation complexity $\eliccvx(\gamma) \geq 2$, since there is no way to draw a $1$-flat (a line, since $p \in \relint{\simplex}$) through $p$ while staying in just one level set on the simplex.

This example also extends to other decision-tree-like properties that do not have an explicit or easily constructed target loss.

>>>>

        \begin{scope}
          \clip (h) -- (m) -- (l) -- (h);
          \foreach \c in {0,0.115,...,1}
          {
            \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
            \pgfmathsetmacro\bb{0.5*\c}
            \coordinate (A) at ($(\aa,1-\aa,0)$);
            \coordinate (B) at ($(1-\aa,\aa,0)$);
            \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
            \draw[myblue,thin] (C) parabola (A);
            \draw[myblue,thin] (C) parabola (B);
          }
          \def\c{0.69}
          \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
          \pgfmathsetmacro\bb{0.5*\c}
          \coordinate (A) at ($(\aa,1-\aa,0)$);
          \coordinate (B) at ($(1-\aa,\aa,0)$);
          \coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
          \coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
          \coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
          %\coordinate (centerdot) at (0.21,0.445, 0.345);
		  \coordinate (centerdot) at (0.3,.4,.3);
          \draw[myblue, very thick] (C) parabola (A);
          \draw[myblue, very thick] (C) parabola (B);
          \draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
          \node at (centerdot) {$\bullet$};
        \end{scope}
        \def\c{0.69}
        \pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
        \pgfmathsetmacro\bb{0.5*\c}
        \node[inner sep=2pt] (levset1) at (0.45,0.55, -0.3) {\vphantom{$\Gamma_{0.69}$}};
        \node (levset) at (0.5,0.5,-.15) {$\Gamma_{0.69}$};
        \draw[-latex,thin] (levset) edge[in=330,out=180] ($(\aa,1-\aa,0)$);
        \draw[-latex,thin] (levset) edge[in=210,out=0] ($(1-\aa,\aa,0)$);
        \placefiglabel{(Q4)}
      \end{tikzpicture}
      \hspace*{-10pt}
      % \\
      % \hline
  \end{tabular}
\caption{Example properties for each quadrant. 
Throughout, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
(Q1,Q2) We cannot fit a $1$-flat (line) through $p$ without leaving the level sets $\gamma_\bot$ and $\gamma_{\text{median}}$, respectively;
Theorem~\ref{thm:convex-flats-inf-dim} implies that there is no $1$-dimensional consistent convex surrogate for either problem.
(Q3) Squared error is a $1$-dimensional convex loss, and indeed it elicits the mean of $Y$, whose level sets are all $1$-flats.
(Q4) The level sets of the variance are curved and cannot fit a $1$-flat; from Theorem~\ref{thm:convex-flats-inf-dim} there is no $1$-dimensional convex surrogate consistent for the variance.}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

Using our $d$-flats framework, we resolve both open questions with new lower bounds in both settings.
% The central observation is that the level sets of $d$-dimensional convex-elicitable properties, if not $d$-flats themselves, are
Our framework clarifies the relationship between $d$-identifiable properties and properties elicited by $d$-dimensional convex losses: the level sets of the former are $d$-flats by definition, while the level sets of the latter are \emph{unions} of $d$-flats by Theorem~\ref{thm:convex-flats-inf-dim}.
A careful examination of the arguments of~\citet{frongillo2020elicitation} and~\citet{dearborn2020indirect} reveals that they largely rely on the containment of $d$-flats in level sets, rather than the full structure of identifiable properties.
As such, although quite subtle in the case of risk measures, the general structure of these previous proofs go through for convex-elicitable properties: since no $d$-flat could be contained in a particular level set, no union of $d$-flats could be either.
Our lower bounds therefore match both of these papers, though we conjecture that the convex consistency dimension is strictly higher \jessie{than convex identification complexity} in some cases.

@@

The problem of estimating a risk or uncertainty measure of $Y$ is of central importance in financial regulation~\cite{acerbi2014back-testing,fissler2016expected,campbell2005review} and robust engineering design\raft{add cites from AFOSR or EnvAI}.
% , and algorithmic fairness~\citep{williamson2019fairness}
Risk measures include the upper confidence bound $\E[Y] + \lambda \sqrt{\Var[Y]}$, or the Conditional Value at Risk (CVaR) defined below in eq.~\eqref{eq:elic-complex-es-3}, in either conditional or unconditional contexts.
Uncertainty measures include the variance, entropy, or norm of the distribution of $Y$.
Risk and uncertainty measures are typically not elicitable, so this problem falls under Quadrant 4.
\citet{frongillo2020elicitation} give prediction dimension lower bounds for a broad class of risk and uncertainty measures, namely Bayes risks.
As stated above, these bounds are with respect to identifiable properties, and bounds for convex surrogates are left as a major open question.

@@

\section{Omitted Examples}\label{app:omitted-examples}
%\subsection{Risk-averse classification (Q2)}
%
%Recall the risk-averse classification from Figure~\ref{fig:quadrants-simplices} (Q2).
%We can bound Convex Consistency Dimension of this classification task by Corollary~\ref{cor:Pcodim-flat-single-val-prop} on the distribution $p = (0.3,0.3,0.4)$, which we can apply since $\gamma(p) = \{$snowy$\}$, suggest that the convex elicitation complexity $\eliccvx(\gamma) \geq 2$, since there is no way to draw a $1$-flat (a line, since $p \in \relint{\simplex}$) through $p$ while staying in just one level set on the simplex.
%
%This example also extends to other decision-tree-like properties that do not have an explicit or easily constructed target loss.
